Galaxy Zoo: Morphological Classifications for 150,000 Galaxies in HST Legacy Imaging

### 2016-05-18 writeup/gz_hubble_data.tex
Use sample labels from Section 2.5 onward

%The creation of 51,861 images from these sources is described in Section
%\ref{ssec:galselect} below.
Single-epoch images from SDSS Stripe~82 were selected using the same criteria
from \citet{wil13}, which required limits of \texttt{petroR90\_r}$ >
3$\arcsec~(where \texttt{petroR90\_r} is the radius containing 90\% of the
$r^\prime$ Petrosian flux) and a magnitude brighter than $m_r < 17.77$. The
images used were three-color composites using the $g^\prime$, $r^\prime$, and
$i^\prime$ filters \citep{nie04}. 21,522 galaxies in SDSS met these criteria.
Co-added images from Stripe~82 were selected from the union of galaxies with
co-added magnitudes brighter than $17.77$~mag, and the galaxies detected in the
single-depth images and matched to a co-add source. This resulted in a total
set of 30,339 images. Of the images in the co-added sample, 5144 (17~percent)
were dimmer than the initial magnitude cut of 17.77.

@@

Images with simulated AGN were classified in the main interface in an identical
manner and evenly distributed with unaltered images of the galaxies.
Classifiers were not explicitly told that the images had been altered during
classification, as the goal was to measure the effect on normal classifications
in as unbiased a manner as possible. Following classification, a classifier
could view a page with additional details about each galaxy they had
classified; where applicable these pages contained further information
regarding image modifications.

@@

In addition to the main \hst{} legacy imaging sample with colored images
described as above, GZH also includes samples of \hst{} galaxy images with
different color prescriptions. In particular, there are 2 sets of 3,927 images
each, drawn from the COSMOS sample; the first has a dramatically reduced color
saturation, and the second has reversed the colors so that the blue and red
filters have exchanged places in the RGB image. The `faded' galaxy set
facilitates measuring of possible variations in classification due to the
presence or absence of color features. The `recolored' set enables an
alternative test of potential color biases in classifications.
% I guess this means we need to say somewhere later that we don't see an effect
% (we don't, right?).  Would also be good to cite here where we have done
% similar things in the past...

@@

%\ref{tbl:catalog_hst}
\item \texttt{main}: \hst{} imaging with RGB colors making use of all available
filters with typical saturation and in correct order, as described in
Section~\ref{ssec:images}. This sample includes AEGIS, COSMOS, GEMS, and
full-depth GOODS (North and South) images. (118,425 galaxies)

%\ref{tbl:catalog_faded}
\item \texttt{faded}: a subset of COSMOS images with very low color saturation.
(3,927 galaxies)

%\ref{tbl:catalog_recolored}
\item \texttt{recolored}: a subset of COSMOS images with red and blue channels
reversed. Note: this sub-sample uses the same galaxies as the \texttt{faded}
sub-sample. (3,927 galaxies)

%\ref{tbl:catalog_goods_shallow}
\item \texttt{goods-shallow}: Images from GOODS-North and GOODS-South, with
color from 2 filters and imaged at 2-epoch depth. (6,144 galaxies)

%\ref{tbl:stripe82_single}
\item \texttt{stripe-82-single}: Single-epoch images from SDSS Stripe~82.
(21,522 galaxies)

%\ref{tbl:stripe82_coadd}
\item \texttt{stripe-82-coadd}: Co-added images from SDSS Stripe~82. Note: this
sub-sample includes all galaxies in the \texttt{stripe-82-single} sub-sample,
with additional sources detected in the deeper imaging. (30,339 galaxies)

%ferengi <-- no table?
\item \texttt{redshifted}: Simulated \hst{} images constructed using SDSS
images and artificially redshifted to between $0.3 \leq z \leq 1$. (288
original galaxies; 3,950 redshifted galaxies)

%\ref{tbl:simulated_agn}
\item \texttt{simulated-agn}: Simulated AGN host galaxies constructed using
\hst{} images and PSFs. (96 original galaxies; 2,961 simulated AGN host
galaxies)

@@

Photometric data for the bulk of the GZH sample were largely drawn from the
tables in \citet{gri12}. This included photometric parameters such as
the fluxes, magnitudes, radii, ellipticities, position angles, and positions
drawn from both \sextractor{} and \galfit. \galfit{} also provided the
parametric \sersic{} index and effective half-light radius for the best-fit
model. All parameters were measured in both bands of the ACS imaging, with the
exception of the single-band COSMOS images.

@@

Photometric and spectroscopic data for the SDSS Stripe~82 galaxies were taken
from the CasJobs DR7 tables. This included $ugriz$ Petrosian magnitudes and
fluxes, as well as the relative de~Vaucouleurs and exponential fits from the
model magnitudes. All redshifts used for SDSS galaxies were spectroscopic.
82.6\% of galaxies in the single-depth images and 65.1\% of galaxies in the
co-add images had a measured DR7 spectroscopic redshift.

@@

The number of independent classifications per subject collected by GZH was on
average higher than GZ1 or GZ2, due to both the increased complexity of the
decision tree and the relative difficulty of classifying images of small and
distant galaxies. Images from the main AEGIS, GEMS, and GOODS data sets had a
median of 122~independent classifications per image. The remainder of the
images either had a later activation date (COSMOS, simulated AGN) or a lower
retirement limit (the low-redshift SDSS Stripe~82 galaxies). Images from these
samples had a median of 46--48~classifications per image
(Figure~\ref{fig:classification_hist}).

@@

The GZH project was launched on 23~Apr~2010 with the inclusion of the AEGIS,
GEMS, GOODS 2-epoch, and SDSS Stripe~82 images. Images from COSMOS and the
simulated AGN were activated in Dec 2010, as well as a small sample of images
from AEGIS, GEMS, and GOODS that were previously excluded from the original
sample due to cuts on blended and/or saturated objects and subsequently
confirmed as classifiable galaxies. The main GZH site collected data until its
replacement, the fourth phase of
Galaxy~Zoo\footnote{\url{http://zoo4.galaxyzoo.org}} (including data from both
the \hst{} CANDELS survey and SDSS DR8) began on 10~Sep~2012. Classifications
for the GOODS 5-epoch images were separately obtained from Mar-Jun 2015 on the
fourth version of the Galaxy~Zoo site. The GZH project had a total of
10,349,357~classifications from 93,898 registered participants. 
% Sum is the total of the GZH SQL table plus the GOODS full-depth imaging in GZ4.

@@

Data products for GZH are split by the type of image being classified.
Table~\ref{tbl:catalog_hst} contains the classifications for the \hst{} images
from the AEGIS, COSMOS, and GEMS surveys, as well as 5-epoch deep imaging from
the GOODS-N and GOODS-S surveys. This contains 118,425~galaxies and is the
primary output from the GZH project. The next two tables have data for a small
subset of 3,927~COSMOS~images that were re-processed to study the effect of
color balance on morphological classification. Table~\ref{tbl:catalog_faded}
has images that are desaturated to minimize the color contrast;
Table~\ref{tbl:catalog_recolored} has images with the red and blue color
channels inverted. Table~\ref{tbl:catalog_goods_shallow} contains data for
6,144~galaxies with 2-epoch images from GOODS. These have been mostly
supplanted in the main table with deeper 5-epoch GOODS imaging; however, there
are 1,683~galaxies in the shallower imaging that were not classified in the
deeper mosaics. This data can also be compared to the counterparts in
Table~\ref{tbl:catalog_hst} to study the effect of depth on morphological
classification. Tables~\ref{tbl:stripe82_single} and \ref{tbl:stripe82_coadd}
contain data for the SDSS Stripe~82 single-depth and co-added images,
respectively, that were classified using the GZH interface and decision tree.
Finally, Table~\ref{tbl:simulated_agn} contains classifications for images with
artificial point sources intended to simulate the effect of a bright AGN (see
\citealt{sim14} for an example).

@@

Given these caveats, the overall distribution of galaxy types is significantly
different than the low-redshift sample classified in SDSS imaging from GZ1 and
GZ2. \citet{lin11} found that elliptical galaxies exceeded spiral galaxies by a
factor of $\sim2:1$ in the main spectroscopic sample if using a plurality vote
criterion (although \citealt{bam09} show that this strongly depends on the
selection method; spirals are the dominant population in a volume-limited
sample at $z<0.088$, for example). The weighted votes for GZH have smooth
galaxies outnumbering disks and clumpy galaxies by a factor of $\sim3:1$,
however. The fraction of objects identified as stars or artifact is also much
higher in the Hubble imaging; by plurality votes, these encompass only
$\sim0.1\%$ of images in SDSS \citep{wil13}, but 6\% of images in GZH.

@@

Morphologies for GEMS galaxies were measured by \citet{hau07}, who used
single-component \sersic{} fits to the F850LP imaging. This analysis used
parameters from the \galfit{} code, which \citet{hau07} evaluated as more
reliable than \gimtwod{} for GEMS due to the ability to fit multiple galaxies
in crowded fields. A 1\arcsec~positional match for the GEMS galaxies in Table~9
in \citet{hau07} to the GZH sample gives 8,846~galaxies in both \citet{hau07}
and GZH. The main morphological parameter in the automated catalog is the
\sersic{} index $n$ defining the radial surface brightness profile.
``Elliptical'' galaxies are selected by $n>2.5$ and ``spirals''  by $n>2.5$.
There is no automatic measurement of irregular or clumpy structure in this
catalog.

@@

GZH used both 5-epoch and 2-epoch sets of data to construct the GOODS set of
images. The 11,157 full depth 5-epoch images are used in the main catalog; the
classifications for the 7,464 shallow depth 2-epoch images are provided as a
supplementary table. This section analyzes the effect of image depth on the
ability of the Galaxy~Zoo classifiers to identify features or disk structure in the images.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.50\textwidth]{figures/full_shallow_p_plot.pdf}
\caption{Distribution of \ffeatures{} for the 4,461 GOODS galaxies with both
shallow (2-epoch) and full-depth (5-epoch) images morphologically classified in
GZH. For most galaxies, the value of \ffeatures{} is consistent ($\Delta
f_\mathrm{features}<0.2$) between depths. Examples of galaxies with sharp
changes in \ffeatures, as well as those with little to no change are shown in
Figures~\ref{fig:shallow_smooth}-\ref{fig:shallow_featured}.}
\label{fig:shallow_vs_full}
\end{center}
\end{figure*}

>>>>

%The creation of 51,861 images from these sources is described in Section
%\ref{ssec:galselect} below.
Single-epoch images from SDSS Stripe~82 were selected using the same criteria
from \citet{wil13}, which required limits of \texttt{petroR90\_r}$ >
3$\arcsec~(where \texttt{petroR90\_r} is the radius containing 90\% of the
$r^\prime$ Petrosian flux) and a magnitude brighter than $m_r < 17.77$. The
images used were three-color composites using the $g^\prime$, $r^\prime$, and
$i^\prime$ filters \citep{nie04}. 21,522 galaxies in SDSS met these criteria.
Co-added images from Stripe~82 were selected from the union of galaxies with
co-added magnitudes brighter than $17.77$~mag, and the galaxies detected in the
\stripe{} images and matched to a co-add source. This resulted in a total
set of 30,339 images. Of the images in the co-added sample, 5144 (17~percent)
were dimmer than the initial magnitude cut of 17.77.

@@

Images with simulated AGN were classified in the interface in an identical
manner and evenly distributed with unaltered images of the galaxies.
Classifiers were not explicitly told that the images had been altered during
classification, as the goal was to measure the effect on normal classifications
in as unbiased a manner as possible. Following classification, a classifier
could view a page with additional details about each galaxy they had
classified; where applicable these pages contained further information
regarding image modifications. \citet{sim14} uses similarly-processed
images for reliable identification of bulgeless galaxies using SDSS data.

@@

In addition to the primary \hst{} legacy imaging sample with colored images
described as above, GZH also includes samples of \hst{} galaxy images with
different color prescriptions. In particular, there are 2 sets of 3,927 images
each, drawn from the COSMOS sample; the first has a dramatically reduced color
saturation, and the second has reversed the colors so that the blue and red
filters have exchanged places in the RGB image. The `faded' galaxy set
facilitates measuring of possible variations in classification due to the
presence or absence of color features. The `recolored' set enables an
alternative test of potential color biases in classifications.
% I guess this means we need to say somewhere later that we don't see an effect
% (we don't, right?).  Would also be good to cite here where we have done
% similar things in the past...

@@

%\ref{tbl:catalog_hst}
\item \main: \hst{} imaging with RGB colors making use of all available
filters with typical saturation and in correct order, as described in
Section~\ref{ssec:images}. This sample includes AEGIS, COSMOS, GEMS, and
full-depth GOODS (North and South) images. (118,425 galaxies)

%\ref{tbl:catalog_faded}
\item \faded: a subset of COSMOS images with very low color saturation.
(3,927 galaxies)

%\ref{tbl:catalog_recolored}
\item \recolored: a subset of COSMOS images with red and blue channels
reversed. Note: this sub-sample uses the same galaxies as the \texttt{faded}
sub-sample. (3,927 galaxies)

%\ref{tbl:catalog_goods_shallow}
\item \goods: Images from GOODS-North and GOODS-South, with
color from 2 filters and imaged at 2-epoch depth. (6,144 galaxies)

%\ref{tbl:stripe82_single}
\item \stripe: Single-epoch images from SDSS Stripe~82.
(21,522 galaxies)

%\ref{tbl:stripe82_coadd}
\item \coadd: Co-added images from SDSS Stripe~82. Note: this
sub-sample includes all galaxies in the \texttt{stripe-82-single} sub-sample,
with additional sources detected in the deeper imaging. (30,339 galaxies)

%ferengi <-- no table?
\item \redshifted: Simulated \hst{} images constructed using SDSS
images and artificially redshifted to between $0.3 \leq z \leq 1$. (288
original galaxies; 3,950 redshifted galaxies)

%\ref{tbl:simulated_agn}
\item \simagn: Simulated AGN host galaxies constructed using
\hst{} images and PSFs. (96 original galaxies; 2,961 simulated AGN host
galaxies)

@@

Photometric data for the bulk of the GZH \main, \faded, \recolored, and \goods{}
samples were largely drawn from the tables in \citet{gri12}. This included
photometric parameters such as the fluxes, magnitudes, radii, ellipticities,
position angles, and positions drawn from both \sextractor{} and \galfit.
\galfit{} also provided the parametric \sersic{} index and effective half-light
radius for the best-fit model. All parameters were measured in both bands of
the ACS imaging, with the exception of the single-band COSMOS images.

@@

Photometric and spectroscopic data for the \stripe{} and \coadd{} galaxies were taken
from the CasJobs DR7 tables. This included $ugriz$ Petrosian magnitudes and
fluxes, as well as the relative de~Vaucouleurs and exponential fits from the
model magnitudes. All redshifts used for SDSS galaxies were spectroscopic.
82.6\% of galaxies in the \stripe{} images and 65.1\% of galaxies in the
\coadd{} images had a measured DR7 spectroscopic redshift.

@@

The number of independent classifications per subject collected by GZH was on
average higher than GZ1 or GZ2, due to both the increased complexity of the
decision tree and the relative difficulty of classifying images of small and
distant galaxies. Images from the \main{} AEGIS, GEMS, and GOODS data sets had
a median of 122~independent classifications per image. The remainder of the
images either had a later activation date (\main{} COSMOS, \simagn) or a lower
retirement limit (\stripe{} and \coadd). Images from these samples had a median
of 46--48~classifications per image (Figure~\ref{fig:classification_hist}).

@@

The GZH project was launched on 23~Apr~2010 with the inclusion of the AEGIS,
GEMS, GOODS 2-epoch, and SDSS Stripe~82 images. Images from COSMOS and the
simulated AGN were activated in Dec 2010, as well as a small sample of images
from AEGIS, GEMS, and GOODS that were previously excluded from the original
sample due to cuts on blended and/or saturated objects and subsequently
confirmed as classifiable galaxies. The GZH site collected data until its
replacement, the fourth phase of
Galaxy~Zoo\footnote{\url{http://zoo4.galaxyzoo.org}} (including data from both
the \hst{} CANDELS survey and SDSS DR8) began on 10~Sep~2012. Classifications
for the GOODS 5-epoch images were separately obtained from Mar-Jun 2015 on the
fourth version of the Galaxy~Zoo site. The GZH project had a total of
10,349,357~classifications from 93,898 registered participants. 
% Sum is the total of the GZH SQL table plus the GOODS full-depth imaging in GZ4.

@@

Data products for GZH are split by the type of image being classified. Each
sample in Section~\ref{ssec:sample_labels} corresponds to the data in
Tables~\ref{tbl:Catalog_hst}-\ref{tbl:simulated_agn}.

@@

Given these caveats, the overall distribution of galaxy types is significantly
different than the low-redshift sample classified in SDSS imaging from GZ1 and
GZ2. \citet{lin11} found that elliptical galaxies exceeded spiral galaxies by a
factor of $\sim2:1$ in their spectroscopic sample if using a plurality vote
criterion (although \citealt{bam09} show that this strongly depends on the
selection method; spirals are the dominant population in a volume-limited
sample at $z<0.088$, for example). The weighted votes for GZH have smooth
galaxies outnumbering disks and clumpy galaxies by a factor of $\sim3:1$,
however. The fraction of objects identified as stars or artifact is also much
higher in the Hubble imaging; by plurality votes, these encompass only
$\sim0.1\%$ of images in SDSS \citep{wil13}, but 6\% of images in GZH.

@@

Morphologies for GEMS galaxies were measured by \citet{hau07}, who used
single-component \sersic{} fits to the F850LP imaging. This analysis used
parameters from the \galfit{} code, which \citet{hau07} evaluated as more
reliable than \gimtwod{} for GEMS due to the ability to fit multiple galaxies
in crowded fields. A 1\arcsec~positional match for the GEMS galaxies in Table~9
in \citet{hau07} to the GZH sample gives 8,846~galaxies in both \citet{hau07}
and GZH. The primary morphological parameter in the automated catalog is the
\sersic{} index $n$ defining the radial surface brightness profile.
``Elliptical'' galaxies are selected by $n>2.5$ and ``spirals''  by $n>2.5$.
There is no automatic measurement of irregular or clumpy structure in this
catalog.

@@

GZH used both 5-epoch and 2-epoch sets of data to construct the GOODS set of
images. The 11,157 full depth 5-epoch images are used in the \main{} catalog; the
classifications for the 7,464 \goods{} images are provided as a
supplementary table. This section analyzes the effect of image depth on the
ability of the Galaxy~Zoo classifiers to identify features or disk structure in the images.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.50\textwidth]{figures/full_shallow_p_plot.pdf}
\caption{Distribution of \ffeatures{} for the 4,461 GOODS galaxies with both
shallow (2-epoch; \goods) and full-depth (5-epoch; \main) images
morphologically classified in GZH. For most galaxies, the value of \ffeatures{}
is consistent ($\Delta f_\mathrm{features}<0.2$) between depths. Examples of
galaxies with sharp changes in \ffeatures, as well as those with little to no
change are shown in
Figures~\ref{fig:shallow_smooth}-\ref{fig:shallow_featured}.}
\label{fig:shallow_vs_full}
\end{center}
\end{figure*}
